---
title: "2018-11-25 20min"
output: html_document
---

---
## Overall goal
1. Identify the list of genes up- and down- regulated upon phosphate starvation
2. Separate these genes by their induction kinetics to early and late group
3. Identify the subset of them that are Pho4-dependent
4. Compare the Pho4-dependent list of genes to previous results based on _pho80??_ background, and discuss any discrepancies
---

---
## 1. Identify the up- and down- regulated genes list upon phosphate starvation
We have a timecourse dataset of wt C.glabrata under phosphate starvation from 20min to 240min. In order to probe the overall sets of genes that can be regulated (either up- or down- regulated) upon phosphate starvation, we need to compare wt C.glabrata strains of distinct time frame of phosphate-depletion with untreated wt C.glabrata strains.

## 1.1 Experiment set-up
Experimental group: wt C.glabrata Phosphate-starvation (20min to 240min)
Control group: wt C.glabrata pre (untreated)

| Experimental group | Control Group | Comment |
| -------- | :---------- | :--------: | ------- |
| wt 20'  | wt pre  | 20 min up- and down- regulated genes upon phosphate starvation |
| wt 30'  | wt pre  | 30 min up- and down- regulated genes upon phosphate starvation |
| wt 45'  | wt pre  | 45 min up- and down- regulated genes upon phosphate starvation |
| wt 60'  | wt pre  | 60 min up- and down- regulated genes upon phosphate starvation |
| wt 90'  | wt pre  | 90 min up- and down- regulated genes upon phosphate starvation |
| wt 120' | wt pre  | 120 min up- and down- regulated genes upon phosphate starvation |
| wt 150' | wt pre  | 150 min up- and down- regulated genes upon phosphate starvation |
| wt 180' | wt pre  | 180 min up- and down- regulated genes upon phosphate starvation |
| wt 240' | wt pre  | 240 min up- and down- regulated genes upon phosphate starvation | 
---

## Prepare data

```{r}
##source("http://www.bioconductor.org/biocLite.R")
##biocLite("VennDiagram")
##biocLite("pheatmap")
##biocLite("NMF")
##biocLite("aroma.light")
library(limma)
library(edgeR)
require(hexbin)
require(ggplot2)       ## for plotting
require(data.table)    ## for fast importing and manipulating data tables
require(cowplot)       ## for some useful defaults in ggplot2
require(RColorBrewer)  ## for plotting
require(gplots)        ## for heatmap plot
library(NMF)           ## for heatmap plot
library(aroma.light)   ## for exploratory analysis
library(knitr)
library(hexbin)
library(VennDiagram)
library(pheatmap)
```


```{r}
1# Input raw data
raw <- fread("../sandbox/Ex009_reads_per_transcript_2017-10-18.txt", check.names = FALSE, stringsAsFactors = FALSE)  ## for the logical arguments, space are needed

# annotation
anno.file <- fread("../data/annotation/C_glabrata_gene_for_mapping_s02-m07-r04.bed")
names(anno.file) <- c("Chr","Start","End","GeneID","Not.Use","Strand","GeneName","Type.of.Gene")
anno <- anno.file[,c("Chr","GeneID","GeneName","Type.of.Gene")]
setkey(anno, "GeneID")
cgToSc <- fread("../data/annotation/C_glabrata_CBS138_S_cerevisiae_orthologs.txt")
names(cgToSc) <- c("cgid","cgname","cgid2","scid","scname","scid2")
setkey(cgToSc, "cgid")

# filter dataset to remove very lowly expressed genes
# 1. examine the distribution of reads for each gene across all samples, to establish a threshold
S <- rowSums(raw[,1:40])
plot(density(log10(S)))
sprintf("The number of genes with less than 40 total counts across 40 samples, i.e. 1 read per sample on average is %d, and those with less than 100 total counts, or 2.5 reads per sample, is %d", sum(S<=40), sum(S<=100))
print("Let's try 40 reads as a cutoff for dropping genes with low or no expression")

## filter dataset
isexpr <- (S <= 40)
use.genes <- grepl("ncRNA|ORF|pseudogene", anno$Type.of.Gene)
use <- (!isexpr & use.genes)
mat <- as.matrix(raw[use, 1:40])
anno.all <- anno; anno <- anno.all[use]

## write out the filtered table (uncomment if changes are made to the filtering steps above)
##write.table(as.data.frame(raw[use]), file = paste("../sandbox/Ex009_reads_per_transcript_filtered_", Sys.Date(), ".txt", sep = ""), quote = FALSE, row.names = FALSE)

rawdata <- fread("../sandbox/Ex009_reads_per_transcript_filtered_2018-12-02.txt", check.names = FALSE, stringsAsFactors = FALSE)
```


```{r}
2# coerce columns to do differentially expressed genes (DEG) analysis
untreated <- as.data.frame.matrix(rawdata[,34:35])
tre.20min <- as.data.frame.matrix(rawdata[,38:39])
tre.30min <- as.data.frame.matrix(rawdata[,2:3])
tre.45min <- as.data.frame.matrix(rawdata[,7:8])
tre.60min <- as.data.frame.matrix(rawdata[,11])
tre.60min2 <- as.data.frame.matrix(rawdata[,13])
tre.60min <- cbind(tre.60min,tre.60min2)
tre.90min <- as.data.frame.matrix(rawdata[,17])
tre.120min <- as.data.frame.matrix(rawdata[,19:20])
tre.150min <- as.data.frame.matrix(rawdata[,23])
tre.180min <- as.data.frame.matrix(rawdata[,26:27])
tre.240min <- as.data.frame.matrix(rawdata[,30])


genename <- as.data.frame.matrix(rawdata[,41])
trial <- cbind(genename,untreated,tre.20min,tre.30min,tre.45min,tre.60min,tre.90min,tre.120min,tre.150min,tre.180min,tre.240min)
write.table(trial,file = "Ex009-wt.txt",sep="\t",row.names=FALSE, quote=FALSE) ## still don't know whether I need sep="\t" here, quote = FALSE means nothing is quoted


```

```{r}
3# normalization
rawdata <- fread("../sandbox/Ex009-wt.txt", check.names = FALSE, stringsAsFactors = FALSE)
y<-DGEList(counts=rawdata[,2:18],genes= anno$GeneID)## [,]means lines or rows of matrix ##DGEList() Creates a DGEList object from a table of counts (rows=features, columns=samples), group indicator for each column, library size (optional) and a table of feature annotation (optional).
y<-calcNormFactors(y) ##TMM normalization, based on the assumption that the majority of genes are not DEGs, calcNormFactors() Calculate normalization factors to scale the raw library sizes.
plotMDS(y)##Plot samples on a two-dimensional scatterplot so that distances on the plot approximate the typical log2 fold changes between the samples.
## calculate log transformed, normalized count
cpm.o <- cpm( y, normalized.lib.sizes = FALSE, log = FALSE) # unnormalized
cpm.o1 <- cpm( y, normalized.lib.sizes = FALSE, log = TRUE) # unnormalized, log transformed
cpm.n1 <- cpm(y,normalized.lib.sizes = TRUE,log= FALSE)#normalized
cpm.n <- cpm(y,normalized.lib.sizes = TRUE,log= TRUE, prior.count=0.5)#normalized, log transformed
boxplot(cpm.o, outline= F)
boxplot(cpm.o1, col="blue",outline= F)
boxplot(cpm.n1, col="red",outline = F)
boxplot(cpm.n, col="green",outline = F)
cpm.n <- cbind(genename,cpm.n)
cpm.n1 <- cbind(genename,cpm.n1)
write.table(cpm.n1,file="TMMnorm-wt-nonlog.txt",row.names=FALSE,sep="\t",quote=F) #normalized
write.table(cpm.n,file="TMMnorm-wt.txt",row.names=FALSE,sep="\t",quote=F) #normalized, log transformed
```

```{r}
4## create design matrix
eset <- read.table("../sandbox/TMMnorm-wt.txt", sep="\t",header = TRUE,row.names=1) ## why we need sep="\t",rownames=1 whether we have the first column for our row names, rather than included into the datafram, which could be no numeric.
eset.pre <- eset[,1:2]
eset.20m <- eset[,3:4]
eset.30m <- eset[,5:6]
eset.45m <- eset[,7:8]
eset.60m <- eset[,9:10]
eset.90m <- eset[,11]
eset.120m <- eset[,12:13]
eset.150m <- eset[,14]
eset.180m <- eset[,15:16]
eset.240m <- eset[,17]


sample.wt <- read.table(text="Sample,Genotype,Timepoint
                         S3,wt,pre
                         S4,wt,pre
                         S7,wt,tre20m
                         s8,wt,tre20m
                         s11,wt,tre30m
                         s12,wt,tre30m
                         s15,wt,tre45m
                         s16,wt,tre45m
                         s19,wt,tre60m
                         s20,wt,tre60m
                         s23,wt,tre90m
                         s25,wt,tre120m
                         s26,wt,tre120m
                         s29,wt,tre150m
                         s31,wt,tre180m
                         s32,wt,tre180m
                         s35,wt,tre240m",sep =",",header =TRUE)
condition <- with(sample.wt,paste(Genotype,Timepoint,sep=".")) ## with(data,expr) construct an environment/list/dataframe, argument expr: expression to evaluate
dsgn.wt <- model.matrix(~0+condition)
colnames(dsgn.wt) <- gsub("condition","",colnames(dsgn.wt))
contrast.matrix <-makeContrasts(
  wtvswt.20min = wt.tre20m - wt.pre,
  wtvswt.30min = wt.tre30m - wt.pre,
  wtvswt.45min = wt.tre45m - wt.pre,
  wtvswt.60min = wt.tre60m - wt.pre,
  wtvswt.90min = wt.tre90m - wt.pre,
  wtvswt.120min = wt.tre120m - wt.pre,
  wtvswt.150min = wt.tre150m - wt.pre,
  wtvswt.180min = wt.tre180m - wt.pre,
  wtvswt.240min = wt.tre240m - wt.pre,
  levels=dsgn.wt)
## makeContrasts(..., contrasts=NULL, levels) 
##Arguments
##...	
##expressions, or character strings which can be parsed to expressions, specifying contrasts
##contrasts=:	character vector specifying contrasts
##levels character vector or factor giving the names of the parameters of which contrasts are desired, or a design matrix or other object with the parameter names as column names.

fit <- lmFit(eset,dsgn.wt)
fit1 <- contrasts.fit(fit,contrast.matrix)
##Given a linear model fit to microarray data, compute estimated coefficients and standard errors for a given set of contrasts.
##Usage
##contrasts.fit(fit, contrasts=NULL, coefficients=NULL)
fit2 <- eBayes(fit1)
sig.wt <- decideTests( fit2, method = "global", lfc = 1 ) # only include genes more than 2 fold induced
choose <- rowSums(sig.wt[,1:2]) >= 1 # select genes that are deemed significant in at least one exp.
test <- coef(fit2)[choose,]
```


```{r}
5# plot the overlap/Venndiagram between the lists
pdf(file="pre-wt-up.pdf",width = 15, height = 7.5)
vennDiagram( fit2[,1:5], include = "up", circle.col = c("turquoise","salmon","orange","blue","yellow") )
dev.off()
pdf(file="pre-wt-down.pdf",width = 15, height = 7.5)
vennDiagram( fit2[,1:5], include = "down", circle.col = c("turquoise","salmon","orange","blue","yellow") )
dev.off()
```


```{r}
6## get DEG list of each Pi-starvation timepoint
dif<-topTable(fit2,coef=c("wtvswt.20min","wtvswt.30min","wtvswt.45min","wtvswt.60min","wtvswt.90min","wtvswt.120min","wtvswt.150min","wtvswt.180min","wtvswt.240min"),n=nrow(fit2),adjust="BH")
genesymbol<-rownames(dif)
dif<-cbind(genesymbol,dif)
write.table(dif,file="probeid.Foldchange-wt.txt",sep='\t',quote=F,row.names=F)
```


```{r}
# 6.sreening diff probeid
dif2<-topTable(fit2,coef=c("wtvswt.20min","wtvswt.30min","wtvswt.45min","wtvswt.60min","wtvswt.90min","wtvswt.120min","wtvswt.150min","wtvswt.180min","wtvswt.240min"),n=nrow(fit2),lfc=log2(4),adjust="BH")
dif3<-dif2[dif2[,"adj.P.Val"]<0.05,]
genesymbol3<-rownames(dif3)
dif3<-cbind(genesymbol3,dif3)
write.table(dif3,file="diff.probeid-FC2-adj.p0.05.txt",sep='\t',quote=F,row.names=F)


## caveat: whether Foldchange=2 is the optimal threshold should be consider carefully in the future, for instance, Tsa1 has fold change less than 2 but a very distinguishable adj.P.value: CAGL0K06259g	wtvswt.20min	wtvswt.30min	wtvswt.45min	wtvswt.60min	wtvswt.90min	wtvswt.120min	wtvswt.150min	wtvswt.180min	wtvswt.240min	AveExpr	F	P.Value	adj.P.Val 1.79713565506261	0.440923936787035	-0.781259550080594	-1.07741129657143	-1.5632439833595	-1.30928068902058	-1.65660221923248	-1.83924038322481	-1.78634622354592	7.49623781767111	64.7451068809915	3.73419673555023e-07	2.72246232075238e-06(adj.p.value)
```

```{r}
# 7.get heatmap of foldchange in all the ORFs
library(pheatmap)
DEG<-read.table("diff.probeid-FC2-adj.p0.05.txt",
                    sep='\t',header=T,row.names=1)
DEG<-DEG[,1:9]
pdf(file="pheatmap-wt-FC.pdf",width=50,height=25)
pheatmap(DEG,
         color=colorRampPalette(c("black","yellow"))(100),
         fontsize_row=0.5,
         fontsize_col=0.5,scale="row",
         border_color=NA,cluster_col = FALSE)
dev.off()

## get out and use 4.probeid to genesymbol.perl to convert all the probe IDs into genesymbol and out put it in the DEG-FC2-0.05-wt.txt
DEG2<-read.table("DEG-FC2-0.05-wt.txt",
                    sep='\t',header=T,row.names=1)
DEG3<-rownames(DEG2)
write.table(DEG3,file="genename-wt.txt",sep='\t',quote=FALSE)
DEG3<-read.table("genename-wt.txt")
DEG2<-DEG2[,1:9]
pdf(file="pheatmap-wt-FCDEGs.pdf",width=10,height=5)
pheatmap(DEG2,
         color=colorRampPalette(c("blue","black","yellow"))(100),
         fontsize_row=0.5,
         fontsize_col=0.5,scale="row",
         border_color=NA,cluster_col = FALSE)
dev.off()
```

```{r}
## 7.1 heatmap of foldchange in DEGs
DEG5<-read.table("DEG-FC2-0.05-wtplus.txt",
                    sep='\t',header=T,row.names=1)
DEG5<-DEG5[,2:10]
thres<-DEG5>=6
DEG5[thres]=6
pheatmap(DEG5,color=colorRampPalette(c("blue","black","yellow"))(100),
         fontsize_row=0.5,
         fontsize_col=0.5,##scale="row",
         border_color=NA,cluster_col = FALSE)
```

```{r}
pdf(file="test.pdf",width=10,height=5)
pheatmap(DEG5,
         color=colorRampPalette(c("blue","black","yellow"))(100),
         fontsize_row=0.5,
         fontsize_col=0.5,##scale="row",
         border_color=NA,cluster_col = FALSE)
dev.off()

```

```{r}
##test for substitution
x=c(1,10,2,8)
i<-x>=6
x[i]=6
pheatmap(x)
```

```{r}
pdf(file="pheatmap-wtplus-FCDEGs.pdf",width=10,height=5)
pheatmap(DEG5,
         color=colorRampPalette(c("blue","black","yellow"))(100),
         fontsize_row=0.5,
         fontsize_col=0.5,scale="row",
         border_color=NA,cluster_col = FALSE)
dev.off()
# another way to plot heatmap
DEG4<-read.table("DEG-wt.txt",
                    sep='\t')
mycol <- colorpanel(1000,"blue","black","yellow")
heatmap.2(as.matrix(DEG2[,1:9]), scale="row",Rowv = TRUE,
   labRow=NA, labCol=colnames(DEG3),
   col=mycol, trace="none", density.info="none", 
   margin=c(4,2), lhei=c(2,10), dendrogram="none")
```



---
##3.Identify the subset of the DEGs that are Pho4-dependent
As a transcription factor, Pho4 translocate into the nucleus only under phosphate-starvation. In order to get the genes that potentially regulated by Pho4, we need to compare the subset of DEGs of WT and del4 strains under phosphate-depletion, and find the subset of DEGs that only regulated by WT strains but not by del4 deleted strains in early and late groups.

## 3.1 Identify the subset of DEGs in del4 groups
Control group:del4 pre (untreated)
Experimental group: del4 treated (20min to 240min)

| Experimental group | Control Group | Comment |
| -------- | :---------- | :--------: | ------- |
| del4 20'  | del4 pre  | 20 min up- and down- regulated genes upon phosphate starvation |
| del4 30'  | del4 pre  | 30 min up- and down- regulated genes upon phosphate starvation |
| del4 45'  | del4 pre  | 45 min up- and down- regulated genes upon phosphate starvation |
| del4 60'  | del4 pre  | 60 min up- and down- regulated genes upon phosphate starvation |
| del4 90'  | del4 pre  | 90 min up- and down- regulated genes upon phosphate starvation |
| del4 120' | del4 pre  | 120 min up- and down- regulated genes upon phosphate starvation |
| del4 150' | del4 pre  | 150 min up- and down- regulated genes upon phosphate starvation |
| del4 180' | del4 pre  | 180 min up- and down- regulated genes upon phosphate starvation |
| del4 240' | del4 pre  | 240 min up- and down- regulated genes upon phosphate starvation | 
---

```{r}
## Input raw data
rawdata <- fread("../sandbox/Ex009_reads_per_transcript_filtered_2018-12-02.txt", check.names = FALSE, stringsAsFactors = FALSE)

# coerce columns of del4 to do differentially expressed genes (DEG) analysis
untreated <- as.data.frame.matrix(rawdata[,36:37])
tre.20min <- as.data.frame.matrix(rawdata[,40])
tre.20min2 <- as.data.frame.matrix(rawdata[,1])
tre.20min <- cbind(tre.20min,tre.20min2)

tre.30min <- as.data.frame.matrix(rawdata[,4])
tre.30min2 <- as.data.frame.matrix(rawdata[,6])
tre.30min <- cbind(tre.30min,tre.30min2)

tre.45min <- as.data.frame.matrix(rawdata[,9:10])
tre.60min <- as.data.frame.matrix(rawdata[,15:16])

tre.90min <- as.data.frame.matrix(rawdata[,18])
tre.120min <- as.data.frame.matrix(rawdata[,21:22])
tre.150min <- as.data.frame.matrix(rawdata[,25])
tre.180min <- as.data.frame.matrix(rawdata[,28:29])
tre.240min <- as.data.frame.matrix(rawdata[,31])


genename <- as.data.frame.matrix(rawdata[,41])
trial <- cbind(genename,untreated,tre.20min,tre.30min,tre.45min,tre.60min,tre.90min,tre.120min,tre.150min,tre.180min,tre.240min)
write.table(trial,file = "Ex009-del4.txt",sep="\t",row.names=FALSE, quote=FALSE) ## still don't know whether I need sep="\t" here, quote = FALSE means nothing is quoted
```

```{r}
## normalization
rawdata <- fread("../sandbox/Ex009-del4.txt", check.names = FALSE, stringsAsFactors = FALSE)
y<-DGEList(counts=rawdata[,2:18],genes= rawdata[,1])## [,]means lines or rows of matrix ##DGEList() Creates a DGEList object from a table of counts (rows=features, columns=samples), group indicator for each column, library size (optional) and a table of feature annotation (optional).
y<-calcNormFactors(y) ##TMM normalization, based on the assumption that the majority of genes are not DEGs, calcNormFactors() Calculate normalization factors to scale the raw library sizes.
plotMDS(y)##Plot samples on a two-dimensional scatterplot so that distances on the plot approximate the typical log2 fold changes between the samples.
## calculate log transformed, normalized count
cpm.o <- cpm( y, normalized.lib.sizes = FALSE, log = FALSE) # unnormalized
cpm.o1 <- cpm( y, normalized.lib.sizes = FALSE, log = TRUE) # unnormalized, log transformed
cpm.n1 <- cpm(y,normalized.lib.sizes = TRUE,log= FALSE)#normalized
cpm.n <- cpm(y,normalized.lib.sizes = TRUE,log= TRUE, prior.count=0.5)#normalized, log transformed
boxplot(cpm.o, outline= F)
boxplot(cpm.o1, col="blue",outline= F)
boxplot(cpm.n1, col="red",outline = F)
boxplot(cpm.n, col="green",outline = F)
cpm.n <- cbind(genename,cpm.n)
cpm.n1 <- cbind(genename,cpm.n1)
write.table(cpm.n1,file="TMMnorm-del4-nonlog.txt",row.names=FALSE,sep="\t",quote=F) #normalized
write.table(cpm.n,file="TMMnorm-del4.txt",row.names=FALSE,sep="\t",quote=F) #normalized, log transformed
```
---