---
title: "2018-11-25 20min"
output: html_document
---
# Identify the subset of the DEGs that are Pho4-dependent
As a transcription factor, Pho4 translocate into the nucleus only under phosphate-starvation. In order to get the genes that potentially regulated by Pho4, we need to compare the subset of DEGs of WT and del4 strains under phosphate-depletion, and find the subset of DEGs that only regulated by WT strains but not by del4 deleted strains in early and late groups.

To call the DEGs of pho4 deleted strains, we have to compare the the differential expression of delpho4 strains under normal or phosphate-depleted conditions. Therefore, our criteria for calling the DEG list is to get genes that differentially expressed in at least one time point treatment with log2(fold change) higher than 2 and adjusted p value (pairwise F-test) less than 0.05. 

## Identify the subset of DEGs in del4 groups
Control group:del4 pre (untreated)
Experimental group: del4 treated (from 20min to 240min)

| Experimental group | Control Group | Comment |
| -------- | :---------- | :--------: | ------- |
| del4 20'  | del4 pre  | 20 min up- and down- regulated genes upon phosphate starvation |
| del4 30'  | del4 pre  | 30 min up- and down- regulated genes upon phosphate starvation |
| del4 45'  | del4 pre  | 45 min up- and down- regulated genes upon phosphate starvation |
| del4 60'  | del4 pre  | 60 min up- and down- regulated genes upon phosphate starvation |
| del4 90'  | del4 pre  | 90 min up- and down- regulated genes upon phosphate starvation |
| del4 120' | del4 pre  | 120 min up- and down- regulated genes upon phosphate starvation |
| del4 150' | del4 pre  | 150 min up- and down- regulated genes upon phosphate starvation |
| del4 180' | del4 pre  | 180 min up- and down- regulated genes upon phosphate starvation |
| del4 240' | del4 pre  | 240 min up- and down- regulated genes upon phosphate starvation | 

# 0.Prepare package
```{r}
source("http://www.bioconductor.org/biocLite.R")
#biocLite("edgeR")
#biocLite("pheatmap")
#biocLite("NMF")
#biocLite("aroma.light")
#install.packages("statmod")
#install.packages("gtools")
#install.packages("hexbin")
#install.packages("data.table")
#install.packages("cowplot")
#install.packages("gplots")
#install.packages("knitr")
#install.packages("cowplot")
#install.packages("RColorBrewer")
#install.packages("dplyr")
#install.packages("miscTools")
library(miscTools)
library(statmod)
library(limma)         ## for differential gene expression analysis
library(edgeR)         ## for DGEList()
library(hexbin)
library(ggplot2)       ## for plotting
library(data.table)    ## for fast importing and manipulating data tables
library(cowplot)       ## for some useful defaults in ggplot2
library(RColorBrewer)  ## for plotting
library(gplots)        ## for heatmap plot
library(NMF)           ## for heatmap plot
library(aroma.light)   ## for exploratory analysis
library(knitr)
library(VennDiagram)
library(pheatmap)
library(gtools)
library(dplyr)
```

# 1. Input raw data
```{r load_data}
raw <- fread("../sandbox/Ex009_reads_per_transcript_2017-10-18.txt", check.names = FALSE, stringsAsFactors = FALSE)
## annotation
anno.file <- fread("../data/annotation/C_glabrata_gene_for_mapping_s02-m07-r04.bed")
names(anno.file) <- c("Chr","Start","End","GeneID","Not.Use","Strand","GeneName","Type.of.Gene") 
anno <- anno.file[,c("Chr","GeneID","GeneName","Type.of.Gene")]
setkey(anno, "GeneID")
cgToSc <- fread("../data/annotation/C_glabrata_CBS138_S_cerevisiae_orthologs.txt")
names(cgToSc) <- c("cgid","cgname","cgid2","scid","scname","scid2")
setkey(cgToSc, "cgid")
## filter dataset to remove very lowly expressed genes
## examine the distribution of reads for each gene across all samples, to establish a threshold
S <- rowSums(raw[,1:40])
Lib.size<-colSums(raw[,1:40])
plot(density(log10(S)))
plot(density(log10(Lib.size)))
sprintf("The number of genes with less than 40 total counts across 40 samples, i.e. 1 read per sample on average is %d, and those with less than 100 total counts, or 2.5 reads per sample, is %d", sum(S<=40), sum(S<=100))
print("Let's try 40 reads as a cutoff for dropping genes with low or no expression")
# filter dataset
isexpr <- (S <= 40)
use.genes <- grepl("ncRNA|ORF|pseudogene", anno$Type.of.Gene)
use <- (!isexpr & use.genes)
mat <- as.matrix(raw[use, 1:40])
anno.all <- anno; anno <- anno.all[use]
```
----
names() sets names for the object 
setkey() sorts a data.table and marks it as sorted, the sorted columns are the key. The columns are sorted in ascending order always
rowSums(),rowMeans(),colSums() form row and column sums and means for numeric arrays(or data frames)
density() computes kernal and bandwidth(sd of kernal), evaluates the density at specific points for data
sprintf() returns a character vector containing a combination of text and variable values; %d conversion by as.character is used for non-character arguments d:interger value
print() returns a character vector
??? isexpr <- (S <= 40)
grepl() searches for mathces to argument pattern with each element, 
   # grepl("1",c(1,2,3)); grepl("ncRNA|ORF|pseudogene",c("ORF|Uncharacterized","pseudogene"))
sub() abd gsub() perform replacement of the fist and all matches respectively
??? (!isexpr&use.genes), ! indicated NOT, & and && indicates AND, | and || indicate OR
??? why u can write raw[use,1:40], use is a matrix with logic values
write out the filtered table (uncomment if changes are made to the filtering steps above)
write.table(as.data.frame(raw[use]), file = paste("../sandbox/Ex009_reads_per_transcript_filtered_", Sys.Date(), ".txt", sep = ""), quote = FALSE, row.names = FALSE)
how u get gene.names into this new table
need to go out and separate one column that cotained all the data to different columns 

# 2. coerce columns to do differentially expressed genes (DEG) analysis(all groups)
```{r}
rawdata <- fread("../sandbox/Ex009_reads_per_transcript_filtered_2018-12-02.txt", check.names = FALSE, stringsAsFactors = FALSE)
sample <- fread("../data/sample_sheet/Ex009_experiment_set_up_20171019.csv",check.names = FALSE, stringsAsFactors = FALSE)
ind.all <- which(!grepl("prime",sample$Sample))
ind.sall <- sample[ind.all,]
ind0 <- match(c("gene.names",ind.sall$Sample),names(rawdata))
ind.all <- rawdata[,ind0,with=F]
```

# 3. normalization (all groups)
```{r}
y.all<-DGEList(counts=ind.all[,2:39],genes=ind.all$gene.names)
y.all<-calcNormFactors(y.all)
plotMDS(cpm(y.all,log=TRUE))
cpm.all <- cpm(y.all,normalized.lib.sizes = TRUE,log= TRUE, prior.count=0.5)
cpm.all1 <- cpm(y.all,normalized.lib.sizes = TRUE,log= FALSE)
boxplot(cpm.all, col="green",outline = F)
boxplot(cpm.all1, col="blue",outline = F)
```
----
[,]means lines or rows of matrix ##DGEList() Creates a DGEList object from a table of counts (rows=features, columns=samples), group indicator for each column, library size (optional) and a table of feature annotation (optional).
TMM normalization, based on the assumption that the majority of genes are not DEGs; calcNormFactors() Calculate normalization factors to scale the raw library sizes.
plotMDS() plot samples on a two-dimensional scatterplot so that distances on the plot approximate the typical log2 fold changes between the samples.
cpm() computes counts per million; rpkm() computes counts per kilobase per million

# 4. create design matrix(all groups)
```{r}
condition.all <- with(ind.sall,paste(Genotype,Timepoint,sep="."))
dsgn.all <- model.matrix(~0+condition.all)
colnames(dsgn.all) <- gsub("condition.all","",colnames(dsgn.all))
contrast.matrix.all <- makeContrasts(
  del4vsdel4.20min = del4.20m - del4.pre,    # Set C
  del4vsdel4.30min = del4.30m - del4.pre,
  del4vsdel4.45min = del4.45m - del4.pre,
  del4vsdel4.60min = del4.60m - del4.pre,
  del4vsdel4.90min = del4.90m - del4.pre,
  del4vsdel4.120min = del4.120m - del4.pre,
  del4vsdel4.150min = del4.150m - del4.pre,
  del4vsdel4.180min = del4.180m - del4.pre,
  del4vsdel4.240min = del4.240m - del4.pre,
  wtvswt.20min = wt.20m - wt.pre,            # Set A
  wtvswt.30min = wt.30m - wt.pre,
  wtvswt.45min = wt.45m - wt.pre,
  wtvswt.60min = wt.60m - wt.pre,
  wtvswt.90min = wt.90m - wt.pre,
  wtvswt.120min = wt.120m - wt.pre,
  wtvswt.150min = wt.150m - wt.pre,
  wtvswt.180min = wt.180m - wt.pre,
  wtvswt.240min = wt.240m - wt.pre,
  wtvswt.del80 = wt.del80 - wt.pre,
  del80vsdel4.del80 = wt.del80 - del4.del80, # Set B
  levels=dsgn.all)
```
----
read.table() reads a file in table format and creates a data frame from it
with(data,expr,...) data to use for constructing an environment
paste() concatenates vectors after converting to character; paste("my","hobby",sep=".")
sub() abd gsub() perform replacement of the fist and all matches respectively
??? model.matrix(~0+condition)
makeContrasts(..., contrasts=NULL, levels) 
??? what is the logic behind model.matrix and makeContrast matrix, why we need to set levels in contrast matrix
expressions, or character strings which can be parsed to expressions, specifying contrasts
contrasts=:	character vector specifying contrasts
levels character vector or factor giving the names of the parameters of which contrasts are desired, or a design matrix or other object with the parameter names as column names.

Question: when we want to compare more than two groups ( not use moderated t-test ), How to use different statistical method to find DEGs? Voom, fit them together with matrix design
functions:
estimateDisp() 
Maximizes the negative binomial likelihood to give the estimate of the different dispersions across all tags.
Think abouT the relationship between normal distribution & Binomial distribution & Negative binomial distribution
Why they assume the dispersion of the data is negative binomial distribution? (variance is higher than mean, compare to Poisson)
estimateDisp()
glmQLFit()
glmQLFTest()
glmTreat()
topTags()

# 5.DEGs Analysis
## 5.1. voom(TMM normalized linear fit of all groups)
```{r}
lcpm.all <- cpm(y.all, normalized.lib.sizes = T, log = F, prior.count = 0.5)
lcpm.all <- cbind(gene.names=ind.all$gene.names,lcpm.all)
v.all <- voom(y.all,dsgn.all,plot=F)
fit.all <- lmFit(v.all,dsgn.all)
fit0 <- contrasts.fit(fit.all,contrast.matrix.all)
fit.all <- eBayes(fit0)
```
## 5.2 Get the DEGs list of 3 compared groups
```{r}
### wt vs. wt no Pi
dif03 <- topTable(fit.all,coef=c("wtvswt.20min","wtvswt.30min","wtvswt.45min","wtvswt.60min","wtvswt.90min","wtvswt.120min","wtvswt.150min","wtvswt.180min","wtvswt.240min"),n=nrow(fit.all),lfc=log2(2),adjust="BH")
dif04 <- dif03[dif03[,"adj.P.Val"]<0.01,]
write.table(dif04,file="all-wt.txt",sep='\t',quote=F,row.names=F)
#### log2(1.8),p<0.01 wt
dif033 <- topTable(fit.all,coef=c("wtvswt.20min","wtvswt.30min","wtvswt.45min","wtvswt.60min","wtvswt.90min","wtvswt.120min","wtvswt.150min","wtvswt.180min","wtvswt.240min"),n=nrow(fit.all),lfc=log2(1.8),adjust="BH")
dif044<-dif033[dif033[,"adj.P.Val"]<0.01,] # Set A
write.table(dif044,file="all-wt-log2(1.8)-p0.01.txt",sep='\t',quote=F,row.names=F)
### del4 vs. del4 no Pi
dif05 <- topTable(fit.all,coef=c("del4vsdel4.20min","del4vsdel4.30min","del4vsdel4.45min","del4vsdel4.60min","del4vsdel4.90min","del4vsdel4.120min","del4vsdel4.150min","del4vsdel4.180min","del4vsdel4.240min"),n=nrow(fit.all),adjust="BH")
write.table(dif05,file="all-del4.txt",sep='\t',quote=F,row.names=F)
### log2(1.8),p<0.01 del4
dif055<-topTable(fit.all,coef=c("del4vsdel4.20min","del4vsdel4.30min","del4vsdel4.45min","del4vsdel4.60min","del4vsdel4.90min","del4vsdel4.120min","del4vsdel4.150min","del4vsdel4.180min","del4vsdel4.240min"),n=nrow(fit.all),lfc=log2(1.8),adjust="BH")
dif066<-dif055[dif055[,"adj.P.Val"]<0.01,] # Pho4 independent targets(PSR) + Pho4 redundant targets(PSR) Set C
write.table(dif066,file="all-del4-log2(1.8)-p0.01.txt",sep='\t',quote=F,row.names=F)
### del80del4 vs. del80
dif07 <- topTable(fit.all,coef="del80vsdel4.del80",n=nrow(fit.all),lfc=log2(1.8),adjust="BH")
dif077 <- dif07[dif07[,"adj.P.Val"]<0.01,] # Set B
write.table(dif077,file="all-del80-log2(1.8)-p0.01.txt",sep='\t',quote=F,row.names=F)
```
## 5.3.1 Get the DEGs independent of Pho4 (TMM-all)
```{r}
###  Compare the genelist of wt and del4 and find their subsets (max FC strategy)
fcdel4.all <- dif066[,1:3]
fcdel4.all <- transform(fcdel4.all,averaged4.23=(fcdel4.all$del4vsdel4.20min+fcdel4.all$del4vsdel4.30min)/2)
fcwt.all <- dif044[,1:3]
fcwt.all <- transform(fcwt.all,averagewt.23=(fcwt.all$wtvswt.20min+fcwt.all$wtvswt.30min)/2)
### filter out downregulated DEGs
fcdel4.all <- subset(fcdel4.all,averaged4.23>1.8,averaged4.23&genes)
fcwt.all <- subset(fcwt.all,averagewt.23>1.8,averagewt.23&genes)
### find the DEGs that are pho4 target candidates
co1 <- match(rownames(fcwt.all),rownames(fcdel4.all));co11 <- sort(co1)
co2 <- fcdel4.all[co11,]
co22 <- match(rownames(co2),rownames(fcwt.all)); 
co222 <- fcwt.all[co22,]
co3 <- cbind(co2,co222$wtvswt.20min,co222$wtvswt.30min,co222$averagewt.23)
colnames(co3) <- c("genes","del4.20min","del4.30min","average.d4","wt.20min","wt.30min","average.wt")
co4 <- transform(co3,minus=co3$average.d4-co3$average.wt)
co5 <- subset(co4,minus>0.6) # 22 genes
co6 <- subset(co4,minus<0.6&minus>0)  # 62 genes
co7 <- subset(co4,minus<0&minus>-0.5) # 71 genes
write.table(co5,file="del4in-log2(0.6).txt",sep='\t',quote=F,row.names=F)
write.table(co6,file="del4in-0tolog2(0.6).txt",sep='\t',quote=F,row.names=F)
write.table(co7,file="del4in-log2(-0.5)to0.txt",sep='\t',quote=F,row.names=F)
```
## 5.3.2 Pho4 independent targets of PSR(set concepts)
```{r}
# Get the set of A,B,C
wt.fc <- dif044[,1:10]
del80.fc <- dif077[,1:2]
del4.fc <- dif066[,1:10]
maxn <- function(n) function(x) order(x, decreasing = TRUE)[n]
wt.fc[,"sec.max"] <- apply(wt.fc[,2:10],1,function(x) x[maxn(2)(x)])
del4.fc[,"sec.max"] <- apply(del4.fc[,2:10],1,function(x) x[maxn(2)(x)])
wt.fc <- subset(wt.fc,wt.fc$sec.max>=0.85)
del4.fc <- subset(del4.fc,del4.fc$sec.max>=0.85)
del80.fc <- subset(del80.fc,del80.fc$logFC>=0.85)
### B∩C 
binterc <- match(del80.fc[,1],del4.fc[,1]); binterc <- sort(binterc)
binterc.upge <- del4.fc[binterc,] # log2=0.85, 150 genes, 114 genes; 41 super sure
write.table(binterc.upge,file="up-BinterC.txt",sep='\t',quote=F,row.names=T)
### (Complementary Set of A)∩(B∩C) (Co+ of Pho4 targets) (all)
reduntar <- binterc.upge[!binterc.upge[,1] %in% wt.fc[,1],] #log2=0.85 9 genes, log2=1.8 10 genes; log2=1 19 genes; log2=0.85 18 gene
###((Complementary Set of B)∩C)
onlyc.upge <- del4.fc[!del4.fc[,1] %in% del80.fc[,1],] #  genes in the onlyCwoB (Pho4 inde PSR genes), includes genes only activated in del4 no pi groups but not activated in wt no Pi groups
write.table(onlyc.upge,file="up-onlyCwoB.txt",sep='\t',quote=F,row.names=T)
### ((Complementary Set of B)∩C)∩A
onlycmod <- wt.fc[wt.fc[,1] %in% onlyc.upge[,1],] # log2=0.85 1094 genes, log2=1.8 376 genes out of 525 genes were also activated in wt no Pi group; log2=1 917 genes
write.table(onlycmod,file="up-Pho4indetargets.txt",sep='\t',quote=F,row.names=T)
```
## 5.4 voom analysis separately
```{r}
### screen DEGs in wt treated groups
dif3 <- topTable(fit.all,coef=c("wtvswt.20min","wtvswt.30min","wtvswt.45min","wtvswt.60min","wtvswt.90min","wtvswt.120min","wtvswt.150min","wtvswt.180min","wtvswt.240min"),n=nrow(fit.wt),lfc=log2(2),adjust="BH")
dif4<-dif3[dif3[,"adj.P.Val"]<0.01,]
write.table(dif4,file="all-wt-DEGs.txt",sep='\t',quote=F,row.names=F)
dif31 <- topTable(fit.all,coef=c("wtvswt.20min","wtvswt.30min","wtvswt.45min","wtvswt.60min","wtvswt.90min","wtvswt.120min","wtvswt.150min","wtvswt.180min","wtvswt.240min"),n=nrow(fit.wt),lfc=log2(4),adjust="BH")
dif41<-dif31[dif31[,"adj.P.Val"]<0.05,]
write.table(dif41,file="all-4FC.txt",sep='\t',quote=F,row.names=F)
### screening overall FC value of del4 groups
dif5<-topTable(fit.all,coef=c("del4vsdel4.20min","del4vsdel4.30min","del4vsdel4.45min","del4vsdel4.60min","del4vsdel4.90min","del4vsdel4.120min","del4vsdel4.150min","del4vsdel4.180min","del4vsdel4.240min"),n=nrow(fit.del4),adjust="BH")
write.table(dif5,file="del4-all2.txt",sep='\t',quote=F,row.names=F)
```
----
Given a linear model fit to microarray data, compute estimated coefficients and standard errors for a given set of contrasts.
contrasts.fit(fit, contrasts=NULL, coefficients=NULL)
# 6. find DEGs dependent on Pho4 (indirect and direct targets)
## 6.1 Using 2nd maximal log2FC value to find Pho4 target candidates (TMM-all)
```{r}
fcw.all <- dif04[,1:10]
### 2nd largest log2FC
maxn <- function(n) function(x) order(x, decreasing = TRUE)[n]
fcw.all[,"sec.max"] <- apply(fcw.all[,2:10],1,function(x) x[maxn(2)(x)])
fcw.all.sec <- as.data.frame(cbind(fcw.all$genes,fcw.all$sec.max));colnames(fcw.all.sec)<-c("genes","sec.max")
write.table(fcw.all.sec,file="sec.max-all-wt.txt",sep='\t',quote=F,row.names=F)
sec.fcw.all <- read.table("../sandbox/sec.max-all-wt.txt",sep='\t',header=T,row.names=1)
### get the max fold change value of all genes del4 no-Pi treated groups
fcd.all <- dif05[,1:10]
fcd.all[,"max"] <- apply(fcd.all[,2:10],1,max)
fcd4.all <- as.data.frame(cbind(fcd.all[,"genes"],fcd.all[,"max"]));colnames(fcd4.all)<-c("genes","max")
write.table(fcd4.all,file="max-all-del4.txt",sep='\t',quote=F,row.names=F)
fc.d4.all <- read.table("../sandbox/max-all-del4.txt",sep='\t',header=T,row.names=1)
### filter out downregulated DEGs
sec.fcw1.all <- subset(sec.fcw.all,sec.max>=0.5,sec.max)
### find the DEGs that are pho4 target candidates
a1 <- match(rownames(sec.fcw1.all),rownames(fc.d4.all))
a2 <- fc.d4.all[a1,]
a3 <- cbind(sec.fcw1.all,a2); colnames(a3) <- c("wt.max","del4.max")
a4 <- transform(a3,minus=a3$wt.max-a3$del4.max)
a33 <- subset(a4,minus>0.6)  # 188 genes
a44 <- subset(a4,minus>0.7)  # 156 genes
a40 <- subset(a4,minus>0.85) # 109 genes
a5 <- subset(a4,minus>1)     # 85 genes
a55 <- subset(a4,minus>1.5)  # 37
a6 <- subset(a4,minus>2)     # 21
a7 <- subset(a4,minus>2.5)   # 10
aa0 <- subset(a4,minus>0.7&minus<1) # 71 genes
aa1 <- subset(a4,minus>1&minus<1.5) # 48 genes
aa2 <- subset(a4,minus>1.5&minus<2) # 16 genes
aa3 <- subset(a4,minus>2&minus<2.5) # 11 genes
# attention! open txt file and align the column names
#write.table(a33,file="TMM-2ndmax-log2(0.6).txt",sep='\t',quote=F,row.names=T) 
#write.table(a44,file="TMM-2ndmax-log2(0.7).txt",sep='\t',quote=F,row.names=T)
#write.table(a40,file="TMM-2ndmax-log2(0.85).txt",sep='\t',quote=F,row.names=T)
#write.table(a4,file="minus-all.txt",sep='\t',quote=F,row.names=T)
# get the early activated and gradually/lately activated Pho4 target candidate list (2nd max TMM-all)
fcw.all <- transform(fcw.all,averagewt.23=(fcw.all$wtvswt.20min+fcw.all$wtvswt.30min)/2,averagewt.late=(fcw.all$wtvswt.180min+fcw.all$wtvswt.240min)/2)
fcw.all <- transform(fcw.all,av.minus=averagewt.late-averagewt.23)
fcw.alsi <- fcw.all[,c("genes","sec.max","av.minus")]
fcw.alsi <- subset(fcw.alsi,sec.max>=0.5,genes&sec.max&av.minus)
se1 <- match(fcw.alsi$genes,fcd4.all$genes)
se2 <- fcd4.all[se1,]
se3 <- cbind(fcw.alsi,se2[,"max"]);colnames(se3) <- c("genes","wt.sec","av.minus","del4.max") # del4.max becomes factors and cannot be calculate
#write.table(se3,file="kinetics-pho4-targets.txt",sep='\t',quote=F,row.names=F)
se4 <- read.table("../sandbox/kinetics-pho4-targets.txt",sep='\t',header=T,row.names=1)
se5 <- transform(se4,ded=se4$wt.sec-se4$del4.max)
se0 <- subset(se5,ded>0.6) # 188 genes
se01 <- subset(se5,ded>0.7) # 156 gebes
se02 <- subset(se5,ded>0.85) # 109 genes
se03 <- subset(se5,ded>1)   # 85 genes
se04 <- subset(se5,ded>1.5) # 37 genes

se77 <- subset(se5,av.minus<1.5) # 1191 gebes find genes that lately activating fold change value is not higher than early activating fold change value
se8 <- subset(se5,av.minus>=1.5) # 232 genes
se9 <- subset(se5,av.minus>=2) # 113 genes
# find early and lately activated genes
se84 <- subset(se8,ded>1.5) # 19 genes lately activated
se774 <- subset(se77,ded>1.5) # 18 genes early activated
```
## 6.2 Using maximum log2FC value to find Pho4 target candidates
```{r}
### get the 2nd and max fold change value of DEGs of wt no-Pi treated groups
fcw <- dif4[,1:10]
### 2nd largest log2FC
maxn <- function(n) function(x) order(x, decreasing = TRUE)[n]
fcw[,"sec.max"] <- apply(fcw[,2:10],1,function(x) x[maxn(3)(x)])
fcwt.sec <- as.data.frame(cbind(fcw$genes,fcw$sec.max));colnames(fcwt.sec)<-c("genes","sec.max")
write.table(fcwt.sec,file="sec.max-wt.txt",sep='\t',quote=F,row.names=F)
sec.fc.wt <- read.table("../sandbox/sec.max-wt.txt",sep='\t',header=T,row.names=1)
### max logFC
fcw[,"max"] <- apply(fcw[,2:10],1,max)
fcwt <- as.data.frame(cbind(fcw$genes,fcw$max));colnames(fcwt)<-c("genes","max")
write.table(fcwt,file="max-wt.txt",sep='\t',quote=F,row.names=F)
fc.wt <- read.table("../sandbox/max-wt.txt",sep='\t',header=T,row.names=1)
### filter out downregulated DEGs
fc.wt1 <- subset(fc.wt,max>=1,max)
sec.fc.wt1 <- subset(sec.fc.wt,sec.max>=1,sec.max)
### get the max fold change value of all genes del4 no-Pi treated groups
fcd <- dif5[,1:10]
fcd[,"max"] <- apply(fcd[,2:10],1,max)
fcd4 <- as.data.frame(cbind(fcd[,"genes"],fcd[,"max"]));colnames(fcd4)<-c("genes","max")
write.table(fcd4,file="max-del4.txt",sep='\t',quote=F,row.names=F)
fc.d4 <- read.table("../sandbox/max-del4.txt",sep='\t',header=T,row.names=1)
### find the DEGs that are pho4 target candidates
m1 <- match(rownames(fc.wt1),rownames(fc.d4))
m2 <- fc.d4[m1,]
m3 <- cbind(fc.wt1,m2); colnames(m3) <- c("wt.max","del4.max")
m4 <- transform(m3,minus=m3$wt.max-m3$del4.max)
m5 <- subset(m4,minus>1)    # 143 genes
m55 <- subset(m4,minus>1.5) # 66
m6 <- subset(m4,minus>2)    # 28
m7 <- subset(m4,minus>2.5)  # 15
m8 <- subset(m4,minus>3)    # 10
write.table(m5,file="pho4-cans-2F.txt",sep='\t',quote=F)
write.table(m6,file="pho4-cans-4F.txt",sep='\t',quote=F)
write.table(m8,file="pho4-cans-8F.txt",sep='\t',quote=F)
### find the DEGs that are pho4 target candidates by 2ndlargest fc
s1 <- match(rownames(sec.fc.wt1),rownames(fc.d4))
s2 <- fc.d4[s1,]
s3 <- cbind(sec.fc.wt1,s2); colnames(s3) <- c("wt.sec.max","del4.max")
s4 <- transform(s3,minus=s3$wt.sec.max-s3$del4.max)
s5 <- subset(s4,minus>1)    # 85 genes
s55 <- subset(s4,minus>1.5) # 37
s6 <- subset(s4,minus>2)    # 20
s7 <- subset(s4,minus>2.5)  # 9
s8 <- subset(s4,minus>3)    # 8
### cluster the DEGs in to different induction levels
l0 <- subset(m4,minus>0.7&minus<1) # 144 genes
l1 <- subset(m4,minus>1&minus<1.5) # 77 genes
l2 <- subset(m4,minus>1.5&minus<2) # 38 genes
l3 <- subset(m4,minus>2&minus<2.5) # 13 genes
ll0 <- subset(s4,minus>0.7&minus<1) # 65 genes
ll1 <- subset(s4,minus>1&minus<1.5) # 48 genes
ll2 <- subset(s4,minus>1.5&minus<2) # 17 genes
ll3 <- subset(s4,minus>2&minus<2.5) # 11 genes
rownames(ll1)
### find the del80 DEGs that are pho4 target candidates
dif.del80 <- read.table("../sandbox/del80-DEGs.txt",sep="\t",header=T,row.names=1)
fc.d8 <- subset(dif.del80,logFC>=2,logFC)
### write.table(fc.d8,file="max-del80.txt",sep='\t',quote=F,row.names=F)
dif.del4del80 <-read.table("../sandbox/del4del80-all.txt",sep="\t",header=T,row.names=1)
dif.del4del80 <- subset(dif.del4del80,logFC>=-1000,logFC)
n1 <- match(rownames(fc.d8),rownames(dif.del4del80))
n2 <- dif.del4del80[n1,]
n3 <- cbind(fc.d8,n2);colnames(n3) <- c("del80","del80del4")
n4 <- transform(n3,minus=n3$del80-n3$del80del4)
n5 <- subset(n4,minus>1)    # 100 genes
n55 <- subset(n4,minus>1.5) # 91 genes
n6 <- subset(n4,minus>2)    # 70 genes
n7 <- subset(n4,minus>2.5)  # 48 genes 
n8 <- subset(n4,minus>3)    # 36 genes
write.table(n5,file="pho4-d8cans-2F.txt",sep='\t',quote=F)
write.table(n6,file="pho4-d8cans-4F.txt",sep='\t',quote=F)
write.table(n8,file="pho4-d8cans-8F.txt",sep='\t',quote=F)
### compare the candidate list between del80 and phosphate-starvation conditions
degm5 <- fread("../sandbox/pho4-cans-2F.txt")
degn5 <- fread("../sandbox/pho4-d8cans-2F.txt")
pho4.fc.minu2 <-intersect(degn5$V1,degm5$V1) #30genes
wtol.minus2 <- setdiff(degm5$V1,degn5$V1) #91 genes
d8ol.minus2 <- setdiff(degn5$V1,degm5$V1) #70 genes

degm6 <- fread("../sandbox/pho4-cans-4F.txt")
degn6 <- fread("../sandbox/pho4-d8cans-4F.txt")
pho4.fc.minu4 <-intersect(degn6$V1,degm6$V1) #11
wtol.minus4 <- setdiff(degm6$V1,degn6$V1) #16 genes
d8ol.minus4 <- setdiff(degn6$V1,degm6$V1) #59 genes

degm8 <- fread("../sandbox/pho4-cans-8F.txt")
degn8 <- fread("../sandbox/pho4-d8cans-8F.txt")
pho4.fc.minu8 <-intersect(degn8$V1,degm8$V1) #5
wtol.minus8 <- setdiff(degm8$V1,degn8$V1) #5 genes
d8ol.minus8 <- setdiff(degn8$V1,degm8$V1) #31 genes
```
## 6.3 (Test!) Using flat slope method to find the definite and direct Pho4 target candidates
```{r}
wt.all <- fread("../sandbox/all-wt.txt")
del4.all <- fread("../sandbox/all-del4.txt")
wt.all <- wt.all[,1:10]
act.wt <- which(wt.all$wtvswt.30min>=0&wt.all$wtvswt.20min>=0)
act.wt <- wt.all[act.wt]
f1 <- match(act.wt$genes,del4.all$genes)
f2 <- del4.all[f1,]
f2 <- f2[,1:10]
my.vec <- c("Timepoint",30,45,60,90,120)
x <- c(f2[2,3:7]); x <- as.numeric(f2[2,3:7]) # do not use rbind because r store vectors in column rather than row; and when combine rows together, the attributes of each elements inside a vector should be same, otherwise it cannot be combined
x <- c(f2[1,3:7]); x <- as.numeric(f2[1,3:7])
y = x
x = as.numeric(my.vec[-1])
lm <- lm(y~x)
summary(lm)
ms <- summary(lm)
coe1 <- ms$coefficients
z1 <- coe1["x","Estimate"]

x <- c(f2[i,2]); x <- as.numeric(f2[i,2])
storage<-numeric(6)
for(i in 1:3){
  x <- c(f2[i,2:3]); x <- as.numeric(f2[i,2:3])
  storage[i] <- x
}
```
----
Loops
1,for loop
for(i in 1:5){
  print(i*2)
# 7. visualize DEGs dynamics in wt and del80 groups
```{r}
## this file contains subfunctions to be sourced into the main analysis for Ex009
sample[,Timepoint:=factor(Timepoint, levels = c("pre","20m","30m","45m","60m","90m","120m","150m","180m","240m"))] ## ,"del80"
sample[,Group:=paste(Genotype,Timepoint, sep=".")]
## normalize all the data with TMM together
### 1. construct count matrix
dge <- DGEList( counts = mat, genes = anno )
### 2. perform TMM normalization
dge <- calcNormFactors(dge)
### 3. use TMM to normalize data
lcpm <- cpm( dge, normalized.lib.sizes = T, log = T, prior.count = 0.5)
rownames(lcpm) <- anno$GeneID
myGenePlot <- function(lc = lcpm, genes = "CAGL0B02475g", sp = sample) {
  # this function takes in the read count matrix (normalized and transformed) and a gene ID
  # and plots the values stratified by genotype and timepoint
  dat <- rbind( lc[genes,sp$Sample] )   # extract the subset of genes used for plotting
  rownames(dat) <- genes
  dt <- data.table(sp[,.(Sample, Genotype, Timepoint)], t(dat))
  dtm <- melt(dt, id = 1:3, variable.name = "GeneID")
  p <- ggplot( dtm, aes( x = Timepoint, y = value, color = Genotype ) )
  print( 
    p + geom_point() + geom_smooth(method = "loess", aes(x = as.numeric(Timepoint), y = value, color = Genotype)) + ylab("log2 count per million") + facet_wrap( ~ GeneID ) + theme(axis.text.x = element_text(angle = 90, size = rel(0.5)))
  )
}
```
## 7.0 PSR Pho4 independent upregulated genes (TM-all&set concepts)
```{r}
myGenePlot(genes = c("CAGL0E02651g","CAGL0F08261g","CAGL0L05434g","CAGL0J04004g","CAGL0A02145g","CAGL0H02101g","CAGL0G02563g","CAGL0B01925g","CAGL0H02563g","CAGL0D04026g","CAGL0J11550g","CAGL0F04719g","CAGL0A02002g","CAGL0L02761g","CAGL0B01078g","CAGL0L05786g","CAGL0D00198g","CAGL0L06094g","CAGL0J03080g","CAGL0I05610g","CAGL0M11902g","CAGL0M12551g")) # Pho4 independent genes but activated during phosphate starvation with an expression >log2(0.6), co5 22 genes
myGenePlot(genes = c("CAGL0C04191g","CAGL0D05082g","CAGL0E05148g","CAGL0I07865g","CAGL0G05544g","CAGL0I05896g","CAGL0F07777g","CAGL0L02827g","CAGL0L02717g"))
myGenePlot(genes = c("CAGL0C02321g","CAGL0L00957g","CAGL0L10582g","CAGL0G09020g","CAGL0M04675g","CAGL0G05335g","CAGL0H08888g","CAGL0K10164g","CAGL0B02563g")) # part of co6
myGenePlot(genes = c("CAGL0J04202g"))#HSP12
### B∩C 41 genes (Part of Pho4 indirect targets(PSR))
myGenePlot(genes = c("CAGL0L04378g","CAGL0A01650g","CAGL0D06600g","CAGL0I07865g","CAGL0M11660g","CAGL0F07777g","CAGL0E01881g","CAGL0L05434g","CAGL0K08844g","CAGL0G02717g","CAGL0H08844g","CAGL0K10164g"))
myGenePlot(genes = c("CAGL0J04004g","CAGL0E01859g","CAGL0G01540g","CAGL0G01738g","CAGL0J02508g","CAGL0H02101g","CAGL0G04433g","CAGL0E05984g","CAGL0K03421g","CAGL0K07546g","CAGL0H08173g","CAGL0C05027g"))
myGenePlot(genes = c("CAGL0K04301g","CAGL0J04202g","CAGL0K03069g","CAGL0E05632g","CAGL0A01606g","CAGL0B02475g","CAGL0F00605g","CAGL0G06050g","CAGL0J02552g","CAGL0G02739g","CAGL0I00726g","CAGL0A00737g"))
myGenePlot(genes = c("CAGL0M03305g","CAGL0E04554g","CAGL0B02882g","CAGL0E01705g","CAGL0K04257g"))
### ((Complementary Set of B)∩C) 525 genes (Pho4 independent targets PSR + genes only activated in del4 no Pi groups)
myGenePlot(genes = c("CAGL0I05390g","CAGL0F05071g","CAGL0M01166g","CAGL0G04081g","CAGL0M12793g","CAGL0G05698g","CAGL0M13541g","CAGL0E03630g","CAGL0C05071g"))
myGenePlot(genes = c("CAGL0I09086g","CAGL0K11275g","CAGL0J09284g","CAGL0C04191g","CAGL0E05148g","CAGL0A01870g","CAGL0C05555g","CAGL0E02651g","CAGL0H07469g"))
myGenePlot(genes = c("CAGL0I05896g","CAGL0M08228g","CAGL0F08261g","CAGL0H02893g","CAGL0E01089g","CAGL0H08063g","CAGL0I05962g","CAGL0L03135g","CAGL0M11308g"))
### ((Complementary Set of B)∩C)∩A 376 genes (Pho4 independent targets PSR)
myGenePlot(genes = c("CAGL0M13541g","CAGL0F05071g","CAGL0G05698g","CAGL0I05390g","CAGL0K11275g","CAGL0M12793g","CAGL0D00990g","CAGL0C04191g","CAGL0A01870g"))
myGenePlot(genes = c("CAGL0F07799g","CAGL0K05137g","CAGL0H07469g","CAGL0E05148g","CAGL0I06138g","CAGL0K07590g","CAGL0E04708g","CAGL0K08162g","CAGL0C05555g"))
myGenePlot(genes = c("CAGL0I05962g","CAGL0C01793g","CAGL0K04235g","CAGL0H02893g","CAGL0I05896g","CAGL0F03113g","CAGL0F07117g","CAGL0M08228g","CAGL0J01331g"))
myGenePlot(genes = c("CAGL0C01771g","CAGL0M13783g","CAGL0J04884g","CAGL0M05665g","CAGL0L06006g","CAGL0M13761g","CAGL0M10439g","CAGL0G05335g","CAGL0L02717g"))
myGenePlot(genes = c("CAGL0L10582g","CAGL0M04675g","CAGL0F03707g","CAGL0J02948g","CAGL0E02651g","CAGL0H08063g","CAGL0I04444g","CAGL0G08624g","CAGL0I01980g"))
```
## 7.00 (TMM-all) early and late activated genes
```{r}
### lately activated se84 fd>1.5
myGenePlot(genes = c("CAGL0A01650g","CAGL0L06644g","CAGL0I00748g","CAGL0M11660g","CAGL0K07524g","CAGL0E05984g","CAGL0G02057g","CAGL0G05984g","CAGL0E00275g","CAGL0E03366g","CAGL0D04400g","CAGL0M03305g","CAGL0L09295g")) #"CAGL0M02915g","CAGL0C00209g","CAGL0A03586g","CAGL0K07678g","CAGL0A03905g","CAGL0G02739g"
### early activated genes se774
myGenePlot(genes = c("CAGL0B02475g","CAGL0K07546g","CAGL0A01243g","CAGL0I04554g","CAGL0D00990g","CAGL0L05456g","CAGL0G06050g","CAGL0D04378g","CAGL0B02453g","CAGL0I03872g","CAGL0K09372g","CAGL0B02970g","CAGL0M02871g","CAGL0B02926g"))#,"CAGL0D01254g","CAGL0L07194g","CAGL0E00803g","CAGL0G03883g"
```
## 7.1 TMM-all
```{r}
### TMM-all >=log2(2.5) FC
myGenePlot(genes = c("CAGL0A01243g","CAGL0B02475g","CAGL0K07546g","CAGL0K07524g","CAGL0I04554g","CAGL0G06050g","CAGL0B02926g","CAGL0D04400g","CAGL0L09295g"))
### TMM-all log2(2)<=FC<=Log2(2.5)
myGenePlot(genes = c("CAGL0L05456g","CAGL0M11660g","CAGL0L06644g","CAGL0M02915g","CAGL0G02057g","CAGL0G05984g","CAGL0C00209g","CAGL0E00803g","CAGL0E03366g","CAGL0M02871g","CAGL0D01254g"))
### TMM-all log2(1.5)<=FC<=Log2(2)
myGenePlot(genes = c("CAGL0A01650g","CAGL0I00748g","CAGL0M12705g","CAGL0D00990g","CAGL0H01177g","CAGL0D04378g","CAGL0B02970g","CAGL0I03872g","CAGL0K07678g","CAGL0E05984g","CAGL0M03305g"))#
### TMM-all log2(1)<=FC<=Log2(1.5)
myGenePlot(genes = c("CAGL0G06952g","CAGL0J07040g","CAGL0C02739g","CAGL0L04378g","CAGL0L12144g","CAGL0H08778g","CAGL0H02387g","CAGL0F07117g","CAGL0F02145g"))
myGenePlot(genes = c("CAGL0I05148g","CAGL0L00737g","CAGL0M12925g","CAGL0K06237g","CAGL0J04884g","CAGL0K12562g","CAGL0E05588g","CAGL0G09262g","CAGL0B03014g","CAGL0J04202g"))
myGenePlot(genes = c("CAGL0C04939g","CAGL0B02453g","CAGL0C01551g","CAGL0G03289g","CAGL0L05258g","CAGL0G03883g","CAGL0B02981g","CAGL0H09218g","CAGL0F02387g"))
myGenePlot(genes = c("CAGL0L06622g","CAGL0J00451g","CAGL0M05137g","CAGL0B04675g","CAGL0F01221g","CAGL0K09372g","CAGL0K04785g","CAGL0F04631g","CAGL0F04191g"))
myGenePlot(genes = c("CAGL0M06347g","CAGL0E00275g","CAGL0E01793g","CAGL0K04257g","CAGL0I04576g","CAGL0A03586g","CAGL0L07194g","CAGL0B02904g","CAGL0M04213g","CAGL0K13002g"))
myGenePlot(genes = c("CAGL0L06644g","CAGL0B02453g","CAGL0I03872g","CAGL0L12144g","CAGL0M12925g","CAGL0K09372g")) # can all be found the minus-all.txt with minus larger than 1
```
```{r}
### TMM-all log2(0.7)<=FC<=Log2(1)
myGenePlot(genes = c("CAGL0C02321g","CAGL0M10351g","CAGL0G01716g","CAGL0K04235g","CAGL0M09449g","CAGL0C00275g","CAGL0H00704g","CAGL0G02101g","CAGL0J02508g"))
myGenePlot(genes = c("CAGL0C05027g","CAGL0I02530g","CAGL0M05951g","CAGL0K04719g","CAGL0M08844g","CAGL0D00946g","CAGL0K10824g","CAGL0E05566g","CAGL0K02893g"))
myGenePlot(genes = c("CAGL0J09900g","CAGL0J01699g","CAGL0H08305g","CAGL0H04719g","CAGL0H02519g","CAGL0M08822g","CAGL0C01397g"))
myGenePlot(genes = c("CAGL0M14091g","CAGL0L04686g","CAGL0A01606g","CAGL0H01606g","CAGL0J06314g","CAGL0K08800g","CAGL0A01284g","CAGL0H06545g","CAGL0G08426g"))
myGenePlot(genes = c("CAGL0E04548g","CAGL0G02695g","CAGL0F08041g","CAGL0J02552g","CAGL0C03311g","CAGL0L07722g","CAGL0A01366g","CAGL0G09251g","CAGL0H06831g"))
myGenePlot(genes = c("CAGL0K01683g","CAGL0J02530g","CAGL0A01111g","CAGL0I02024g","CAGL0K03201g","CAGL0L11990g","CAGL0H05995g","CAGL0B02948g","CAGL0I04004g"))
myGenePlot(genes = c("CAGL0I00726g","CAGL0E06182g","CAGL0G08195g","CAGL0K00170g","CAGL0E01177g","CAGL0G06182g","CAGL0K08036r"))
```
```{r}
#### log2FC between 1.62 to 2
myGenePlot(genes = c("CAGL0K11275g","CAGL0C02321g","CAGL0M13541g","CAGL0F00605g","CAGL0M10351g","CAGL0D06600g","CAGL0G01716g","CAGL0K07458g"))
myGenePlot(genes = c("CAGL0C01661g","CAGL0I07865g","CAGL0L04664g","CAGL0K12034g","CAGL0E06072g","CAGL0K07590g","CAGL0J01331g","CAGL0J04004g"))
myGenePlot(genes = c("CAGL0C01793g","CAGL0D03520g","CAGL0G01738g","CAGL0M08552g","CAGL0F08107g","CAGL0J08954g","CAGL0I01496g","CAGL0A01870g"))
myGenePlot(genes = c("CAGL0G02849g","CAGL0M05665g","CAGL0K06259g","CAGL0F00869g","CAGL0M05951g","CAGL0D03542g","CAGL0L01265g","CAGL0F04807g"))
```
```{r}
#### log2FC between 2 to 2.8 max log2FC
myGenePlot(genes = c("CAGL0J07040g","CAGL0C02739g","CAGL0L04378g","CAGL0K04235g","CAGL0L12144g","CAGL0H08778g","CAGL0H02387g","CAGL0F07117g","CAGL0M09449g"))
myGenePlot(genes = c("CAGL0B02860g","CAGL0C00275g","CAGL0H00704g","CAGL0F02145g","CAGL0I05148g","CAGL0G02101g","CAGL0J02508g","CAGL0K06237g","CAGL0D00946g"))
myGenePlot(genes = c("CAGL0K12562g","CAGL0C05027g","CAGL0E05588g","CAGL0K04037g","CAGL0I02530g","CAGL0B03014g","CAGL0M08844g","CAGL0K10824g","CAGL0E05566g"))
myGenePlot(genes = c("CAGL0J09900g","CAGL0H04719g","CAGL0H02519g","CAGL0H09218g","CAGL0M08822g","CAGL0C01397g","CAGL0F02387g","CAGL0L06622g","CAGL0E04884g"))
myGenePlot(genes = c("CAGL0M14091g","CAGL0A01606g","CAGL0H01606g","CAGL0K03459g","CAGL0M05137g","CAGL0B04675g","CAGL0A01284g","CAGL0F01221g","CAGL0H06545g"))
#### log2FC between 2 to 2.8 2ndmax log2FC
myGenePlot(genes = c("CAGL0J07040g","CAGL0C02739g","CAGL0L04378g","CAGL0K04235g","CAGL0L12144g","CAGL0H08778g","CAGL0H02387g","CAGL0F07117g","CAGL0C00275g"))
myGenePlot(genes = c("CAGL0F02145g","CAGL0I05148g","CAGL0L00737g","CAGL0M12925g","CAGL0K06237g","CAGL0K12562g","CAGL0C05027g","CAGL0E05588g","CAGL0G09262g"))
myGenePlot(genes = c("CAGL0B03014g","CAGL0J04202g","CAGL0C01551g","CAGL0G03289g","CAGL0K10824g","CAGL0L05258g","CAGL0B02981g","CAGL0H09218g","CAGL0C01397g"))
myGenePlot(genes = c("CAGL0F02387g","CAGL0L06622g","CAGL0J00451g","CAGL0M05137g","CAGL0B04675g","CAGL0F01221g","CAGL0K04785g","CAGL0F04631g","CAGL0F04191g"))
myGenePlot(genes = c("CAGL0M06347g","CAGL0E00275g","CAGL0E01793g","CAGL0H06831g","CAGL0K01683g","CAGL0K03201g","CAGL0K04257g","CAGL0I04576g","CAGL0A03586g"))
```
```{r}
#### log2FC between 2.8 to 4
myGenePlot(genes = c("CAGL0A01650g","CAGL0G06952g","CAGL0I00748g","CAGL0M12705g","CAGL0L00737g","CAGL0M12925g","CAGL0J04884g","CAGL0G09262g"))
myGenePlot(genes = c("CAGL0H01177g","CAGL0J04202g","CAGL0C04939g","CAGL0B02453g","CAGL0C01551g","CAGL0G03289g","CAGL0L05258g","CAGL0G03883g"))
myGenePlot(genes = c("CAGL0D04378g","CAGL0B02981g","CAGL0L04686g","CAGL0J00451g","CAGL0K09372g","CAGL0F04191g","CAGL0E00275g","CAGL0K04257g"))
myGenePlot(genes = c("CAGL0L11990g","CAGL0A03586g","CAGL0L07194g","CAGL0K00170g","CAGL0B02904g","CAGL0K13002g"))
```
## 7.2 visualize the most significant log2FC via max log2FC method
```{r}
## 8 FC difference
myGenePlot(genes = c("CAGL0A01243g","CAGL0B02475g","CAGL0K07546g","CAGL0K07524g","CAGL0I04554g","CAGL0G06050g","CAGL0B02926g","CAGL0C00209g","CAGL0D04400g","CAGL0L09295g"))
## EPA gene difference
myGenePlot(genes = c("CAGL0E06644g","CAGL0C00110g","CAGL0C05643g","CAGL0C00847g","CAGL0A01366g","CAGL0A01284g","CAGL0L13299g","CAGL0M00132g","CAGL0L13332g","CAGL0J11968g","CAGL0E00275g","CAGL0K00170g","CAGL0I00220g"))
## EPA1,6,7,20
myGenePlot(genes = c("CAGL0E06644g","CAGL0C00110g","CAGL0C05643g","CAGL0E00275g"))
## Bmt1,2,3,4,5,6,7
myGenePlot(genes = c("CAGL0D00286g","CAGL0B02882g","CAGL0B02926g","CAGL0B02948g","CAGL0B02970g","CAGL0B02904g","CAGL0K12980g"))
## intersection of pho4 dependent DEGs in del80 and no-Pi conditions
### 8 FC difference
myGenePlot(genes = c("CAGL0A01243g","CAGL0B02475g","CAGL0K07546g","CAGL0B02926g","CAGL0G06050g"))
### 4 FC difference
myGenePlot(genes = c("CAGL0M11660g","CAGL0A01243g","CAGL0B02475g","CAGL0K07546g","CAGL0B02926g","CAGL0G06050g","CAGL0E05984g","CAGL0M02915g","CAGL0E03366g","CAGL0L05456g","CAGL0C00209g"))
## genes only activated in pho80 deleted groups (3list)
myGenePlot(genes = c("CAGL0K10164g","CAGL0E04554g","CAGL0I10791g","CAGL0G01540g","CAGL0K04301g","CAGL0F00649g","CAGL0M13189g"))

```
## 7.3 visualize mildly induced DEGs according to clustered heatmap of lfc=log2(4)&P<0.05 of wt DEGs, FC value was set to be saturated to 4.
```{r}
###selected mildly induced from heatmap
#### >=log2(7)
myGenePlot(genes = c("CAGL0H05379g","CAGL0K01683g","CAGL0L06644g","CAGL0B02453g","CAGL0K09878g","CAGL0H05379g","CAGL0K01683g"))
#### log2(4)<= <=log2(7)
myGenePlot(genes = c("CAGL0L12144g","CAGL0G08426g","CAGL0I03872g","CAGL0J11176g","CAGL0M12925g","CAGL0G01716g"))
#### <=log2(2)
myGenePlot(genes = c("CAGL0K04785g","CAGL0K08036r","CAGL0F00341g","CAGL0L07766g","CAGL0A03586g","CAGL0E01749g"))
#### del80 is very different to wt,might still be activated after 4 hours
myGenePlot(genes = c("CAGL0B01100g","CAGL0D01276g","CAGL0K04037g","CAGL0I09284g"))
###del80 is very different to wt, do not know how to explain
myGenePlot(genes = c("CAGL0M13717g","CAGL0I10098g","CAGL0C04719g","CAGL0D01837g","CAGL0L10142g","CAGL0J07953r"))
```
## 7.4 the overall Pho4 targets (Set concepts)(PSR+other responses)
```{r}
### A∪B-((Complementary Set of B)∩C)∩A log2=0.85 370+25=395 genes; log2=1 252+23=275 genes; log2=1.8 252+16=268 genes;
del80.pho4 <- del80.fc[!del80.fc[,1] %in% onlycmod[,1],] #log2=0.85 182 genes; log2=1 149 genes; ; log2=1.8 60 genes
wt.pho4 <- wt.fc[!wt.fc[,1] %in% onlycmod[,1],] #log2=0.85 370 genes; log2=1 363 genes; log2=1.8 252 genes;\
test1 <- del80.fc[!del80.pho4[,1] %in% wt.pho4[,1],] #log2=0.85 25 genes, 15 genes; log2=1 23 genes
### Pho4 overall targets (PSR+Other) = genes from test1&wt.pho4
### Pho4 targets (PSR) = genes from wt.pho4
### Pho4 targets log2=1
myGenePlot(genes = c("CAGL0B02475g","CAGL0A01650g","CAGL0G06952g","CAGL0K07546g","CAGL0J07040g","CAGL0A01243g","CAGL0C02739g","CAGL0M10351g","CAGL0A02321g"))
myGenePlot(genes = c("CAGL0K12254g","CAGL0C01661g","CAGL0D06600g","CAGL0I04554g","CAGL0L00737g","CAGL0L05456g","CAGL0L04378g","CAGL0L04598g","CAGL0C02321g"))
myGenePlot(genes = c("CAGL0F03685g","CAGL0F02145g","CAGL0E05588g","CAGL0I07865g","CAGL0G05544g","CAGL0D05082g","CAGL0F07777g","CAGL0L12144g","CAGL0J06050g"))
myGenePlot(genes = c("CAGL0C04543g","CAGL0K07458g","CAGL0K11968g","CAGL0D03520g","CAGL0H02387g","CAGL0M12705g","CAGL0G09977g","CAGL0L05016g","CAGL0L06644g"))
myGenePlot(genes = c("CAGL0D03542g","CAGL0E01881g","CAGL0I00748g","CAGL0J08547g","CAGL0C04939g","CAGL0D04070g","CAGL0H08844g","CAGL0L02827g","CAGL0J11176g"))
### Pho4 targets log2=1.8
myGenePlot(genes = c("CAGL0B02475g","CAGL0A01650g","CAGL0G06952g","CAGL0K07546g","CAGL0J07040g","CAGL0G01716g","CAGL0A01243g","CAGL0M09449g","CAGL0C02739g"))
myGenePlot(genes = c("CAGL0M10351g","CAGL0K12254g","CAGL0L06204g","CAGL0D06600g","CAGL0I04554g","CAGL0L00737g","CAGL0L05456g","CAGL0L04378g","CAGL0L04598g"))
myGenePlot(genes = c("CAGL0C02321g","CAGL0F02145g","CAGL0E05588g","CAGL0I07865g","CAGL0G05544g","CAGL0D05082g","CAGL0F07777g","CAGL0J06050g","CAGL0C04543g"))
myGenePlot(genes = c("CAGL0K07458g","CAGL0K01221g","CAGL0H02387g","CAGL0M12705g","CAGL0G09977g","CAGL0L05016g","CAGL0L06644g","CAGL0E01881g","CAGL0I00748g"))
myGenePlot(genes = c("CAGL0J08547g","CAGL0C04939g","CAGL0H08844g","CAGL0L02827g","CAGL0K04037g","CAGL0J11176g","CAGL0M06259g","CAGL0M11660g","CAGL0B02563g"))
#### test1 log2=1.8
myGenePlot(genes = c("CAGL0K08844g","CAGL0K05390g","CAGL0M02629g","CAGL0C00110g","CAGL0D05170g","CAGL0L03267g","CAGL0E05632g","CAGL0J06028g","CAGL0L02321g","CAGL0E06292g","CAGL0L03157g","CAGL0E01705g","CAGL0L08547g","CAGL0C05643g","CAGL0J00847g","CAGL0I00286g"))
```
# 7.5. Find Pho4 redundant targets(not mainly regulated by Pho4)
```{r}


```
# 8.get heatmap of foldchange in DEGs
```{r}
dif.wt<-read.table("wt-DEGs.txt",
                    sep='\t',header=T,row.names=1)
dif.wt<-dif.wt[,1:9]
pheatmap(dif.wt,
         color=colorRampPalette(c("blue","black","yellow"))(100),
         fontsize_row=0.5,
         fontsize_col=0.5,scale="none",
         border_color=NA,cluster_col = FALSE)

dif.wt2<-read.table("wt-4FC.txt",
                    sep='\t',header=T,row.names=1)
dif.wt2<-dif.wt2[,1:9]
pheatmap(dif.wt2,
         color=colorRampPalette(c("blue","black","yellow"))(100),
         fontsize_row=0.5,
         fontsize_col=0.5,scale="none",
         border_color=NA,cluster_col = FALSE)

dif.del4 <-read.table("del4-DEGs.txt",
                    sep='\t',header=T,row.names=1)
dif.del4<-dif.del4[,1:9]
pheatmap(dif.del4,
         color=colorRampPalette(c("blue","black","yellow"))(100),
         fontsize_row=0.5,
         fontsize_col=0.5,scale="none",
         border_color=NA,cluster_col = FALSE)
```

## 8.2 use heatmap to visualize genes with lower foldchange
```{r}
thres.wt <- dif.wt>=4
dif.wt[thres.wt]=4
pheatmap(dif.wt,
         color=colorRampPalette(c("blue","black","yellow"))(100),
         fontsize_row=0.5,
         fontsize_col=0.5,scale="none",
         border_color=NA,cluster_col = FALSE)

thres.wt2 <- dif.wt2>=4
dif.wt2[thres.wt2]=4
pheatmap(dif.wt2,
         color=colorRampPalette(c("blue","black","yellow"))(100),
         fontsize_row=0.5,
         fontsize_col=0.5,scale="none",
         border_color=NA,cluster_col = FALSE)

thres.del4 <- dif.del4>=4
dif.del4[thres.del4]=4
pheatmap(dif.del4,
         color=colorRampPalette(c("blue","black","yellow"))(100),
         fontsize_row=0.5,
         fontsize_col=0.5,scale="none",
         border_color=NA,cluster_col = FALSE)
```

```{r}
pdf(file="wt-4FCs-thre.pdf",width=50,height=25)
pheatmap(dif.wt2,
         color=colorRampPalette(c("blue","black","yellow"))(100),
         fontsize_row=0.5,
         fontsize_col=0.5,scale="none",
         border_color=NA,cluster_col = FALSE)
dev.off()
```

# 9 Compare pho80-deleted background, CHIPseq data and my 2nd-max pho4 candidates list
## 9.1 log2(0.6)
```{r}
chip.data <- fread(file="../sandbox/elife-25157-fig5-data1-v2.csv", sep= ",",sep2=",",header = T)
Rseq80.data <- fread(file="../sandbox/elife-25157-fig5-data2-v2.csv", sep= ",",sep2=",",header = T)
my.list <- fread(file="TMM-2ndmax-log2(0.6).txt", sep= ",",sep2=",",header = T) # 188 genes
## Compare RNAseq gene list (my list to Bin's list)
Rseqm <- match(Rseq80.data$GeneID,my.list$genes) # 37 out of 80 genes in Bin's RNAseq list could match to my list 
Rseqm <- sort(Rseqm)
Rseqm1 <- my.list[Rseqm,]
## ChIP - Bin - My
Chipse <- match(chip.data$START,Rseq80.data$START) # 70 out of 100 Chipseq data could match to Rseq80
Chipse <- sort(Chipse)
Chipse1 <- Rseq80.data[Chipse,]
comatch <- match(Chipse1$GeneID,my.list$genes) # 31 out of 70 could match to my list
comatch <- sort(comatch)
comatch1 <- my.list[comatch]
## My - Bin - ChIP
mytb <- match(my.list$genes,Rseq80.data$GeneID) # 37 genes in my list coud match to Bin's list 
mytb <- sort(mytb)
mytb1 <- Rseq80.data[mytb,]
mytbtc <- match(mytb1$START,chip.data$START) # all 37 genes could match to both list
write.table(mytb1,file="3listSub-log2(0.6).txt",sep='\t',quote=F,row.names=F)
```
## 9.2 log2(0.7)
```{r}
chip.data <- fread(file="../sandbox/elife-25157-fig5-data1-v2.csv", sep= ",",sep2=",",header = T)
Rseq80.data <- fread(file="../sandbox/elife-25157-fig5-data2-v2.csv", sep= ",",sep2=",",header = T)
my.list <- fread(file="TMM-2ndmax-log2(0.7).txt", sep= ",",sep2=",",header = T) # 156 genes
## Compare RNAseq gene list (my list to Bin's list)
Rseqm <- match(Rseq80.data$GeneID,my.list$genes) # 36 out of 80 genes in Bin's RNAseq list could match to my list 
Rseqm <- sort(Rseqm)
Rseqm1 <- my.list[Rseqm,]
## ChIP - Bin - My
Chipse <- match(chip.data$START,Rseq80.data$START) # 70 out of 100 Chipseq data could match to Rseq80
Chipse <- sort(Chipse)
Chipse1 <- Rseq80.data[Chipse,]
comatch <- match(Chipse1$GeneID,my.list$genes) # 30 out of 70 could match to my list
comatch <- sort(comatch)
comatch1 <- my.list[comatch]
## My - Bin - ChIP
mytb <- match(my.list$genes,Rseq80.data$GeneID) # 36 genes in my list coud match to Bin's list 
mytb <- sort(mytb)
mytb1 <- Rseq80.data[mytb,]
mytbtc <- match(mytb1$START,chip.data$START) # all 36 genes could match to both list
write.table(mytb1,file="3listSub-log2(0.7).txt",sep='\t',quote=F,row.names=F)
```
## 9.3 log2(0.85)
```{r}
chip.data <- fread(file="../sandbox/elife-25157-fig5-data1-v2.csv", sep= ",",sep2=",",header = T)
Rseq80.data <- fread(file="../sandbox/elife-25157-fig5-data2-v2.csv", sep= ",",sep2=",",header = T)
my.list <- fread(file="TMM-2ndmax-log2(0.85).txt", sep= ",",sep2=",",header = T) # 156 genes
## Compare RNAseq gene list (my list to Bin's list)
Rseqm <- match(Rseq80.data$GeneID,my.list$genes) # 28 out of 80 genes in Bin's RNAseq list could match to my list 
Rseqm <- sort(Rseqm)
Rseqm1 <- my.list[Rseqm,]
## ChIP - Bin - My
Chipse <- match(chip.data$START,Rseq80.data$START) # 70 out of 100 Chipseq data could match to Rseq80
Chipse <- sort(Chipse)
Chipse1 <- Rseq80.data[Chipse,]
comatch <- match(Chipse1$GeneID,my.list$genes) # 23 out of 70 could match to my list
comatch <- sort(comatch)
comatch1 <- my.list[comatch]
## My - Bin - ChIP
mytb <- match(my.list$genes,Rseq80.data$GeneID) # 28 genes in my list coud match to Bin's list 
mytb <- sort(mytb)
mytb1 <- Rseq80.data[mytb,]
mytbtc <- match(mytb1$START,chip.data$START) # all 36 genes could match to both list
write.table(mytb1,file="3listSub-log2(0.85).txt",sep='\t',quote=F,row.names=F)
```